---
title: "Meta_analysis_keep_500genes_comp1"
output: html_document
date: "2024-06-13"
author: "LANON Alizée"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,     # Suppress warnings from R code
  message = FALSE     
)
```
## Download dataset
```{r, message=FALSE}
################################################################################################################################################
#### STEPNIAK
data_stepniak <- read.table("../../../Glioma/Stepniak/RNAseq/nextflow/star_salmon/salmon.merged.gene_counts.tsv", header = TRUE)
pdata_stepniak <- read.csv("pdata_DA_complete.csv", sep =";", header = TRUE)
row.names(pdata_stepniak) <- pdata_stepniak$Sample
#--> Count matrix
mrna_raw_counts_stepniak <- data_stepniak
row.names(mrna_raw_counts_stepniak) <- mrna_raw_counts_stepniak$gene_id
mrna_raw_counts_stepniak <- mrna_raw_counts_stepniak[,-c(1:2)]
mrna_raw_counts_stepniak <- mrna_raw_counts_stepniak[,c(1,4,3,6,5,7)]
#--> Order and match
mrna_raw_counts_stepniak <- mrna_raw_counts_stepniak[,match(pdata_stepniak$Sample, colnames(mrna_raw_counts_stepniak), nomatch = 0)]
pdata_stepniak <- pdata_stepniak[pdata_stepniak$Sample %in% colnames(mrna_raw_counts_stepniak),]
################################################################################################################################################
#### JIGNA
data_jigna <- read.table("salmon.merged.gene_counts_Jigna.tsv", header = TRUE)
pdata_jigna <- read.csv("pdata_Jigna.csv", sep =";", header = TRUE)
row.names(pdata_jigna) <- pdata_jigna$Sample
#--> Count matrix
mrna_raw_counts_jigna <- data_jigna
row.names(mrna_raw_counts_jigna) <- mrna_raw_counts_jigna$gene_id
mrna_raw_counts_jigna <- mrna_raw_counts_jigna[,-c(1:2)]
mrna_raw_counts_jigna <- mrna_raw_counts_jigna[,c(1,3,5,7,9,11,13,15)]
#--> Order and match
mrna_raw_counts_jigna <- mrna_raw_counts_jigna[,match(pdata_jigna$Sample, colnames(mrna_raw_counts_jigna), nomatch = 0)]
pdata_jigna <- pdata_jigna[pdata_jigna$Sample %in% colnames(mrna_raw_counts_jigna),]
################################################################################################################################################
```
## COMBIBNE THE 2 DATAFRAMES AND FILTERING
```{r, message=FALSE}
library(HTSFilter)
common_lines <- intersect(rownames(mrna_raw_counts_jigna), rownames(mrna_raw_counts_stepniak))

mrna_raw_counts_jigna_common <- mrna_raw_counts_jigna[common_lines,]
mrna_raw_counts_stepniak_common <- mrna_raw_counts_stepniak[common_lines,]
combined_count_mrna <- cbind(mrna_raw_counts_jigna_common,mrna_raw_counts_stepniak_common)

Y_pdata <- read.csv("pdata.csv", sep =";", header = TRUE)

mrna_filter_common <- HTSFilter(round(combined_count_mrna),
                         Y_pdata$Group,
                         s.min=1, s.max=200)

mrna_filter_common <- mrna_filter_common$filteredData
```
## SEAPARATION AFTER FILTERING TO NORMALIZE INDEPENDENTLY
```{r, message=FALSE}
mrna_filter_stepniak <- mrna_filter_common[,9:14]
mrna_filter_jigna <- mrna_filter_common[,1:8]
```
## STEPNIAK DATA NORMALIZATION
```{r, message=FALSE}
library(mixOmics)

#--> Feature annotation
mrna_gene_stepniak <- read.table("../../../Glioma/Stepniak/RNAseq/nextflow/star_salmon/salmon.merged.gene_counts.tsv", header = TRUE)[,c(1:2)]

## Normalisation
library(RUVSeq)
library(reshape)
library(EDASeq)

#--> edgeR TMM normalization
library(edgeR)

tmm_set_stepniak <-  mrna_filter_stepniak

pdata_tmm_stepniak <- pdata_stepniak
design <- cbind(model.matrix(~ 0 + pdata_tmm_stepniak$Group))
colnames(design) <- gsub(".*\\$|)", "", colnames(design))

tmm_set_stepniak <- DGEList(counts = tmm_set_stepniak,
                   genes = mrna_gene_stepniak[match(
                     row.names(mrna_filter_stepniak), 
                     mrna_gene_stepniak$gene_id),],
                   group = pdata_tmm_stepniak$Group)

tmm_set_stepniak <- calcNormFactors(tmm_set_stepniak)

#--> Estimate dispersion and fit the model
tmm_set_stepniak <- estimateGLMRobustDisp(tmm_set_stepniak, design)

##### Count per million
tmm_cpm_stepniak <- cpm(tmm_set_stepniak, normalized.lib.sizes = TRUE, log = TRUE)

# --> Transpose the df to have the sample in lines
t_mrna_norm_stepniak <- as.data.frame(t(tmm_cpm_stepniak))

```
## JIGNA DATA NORMALIZATION
```{r, message=FALSE}
#--> Feature annotation
mrna_gene_jigna <- read.table("salmon.merged.gene_counts_Jigna.tsv", header = TRUE)[,c(1:2)]

###### Normalization ######
library(RUVSeq)
library(reshape)
library(EDASeq)

#--> edgeR TMM normalization
library(edgeR)

tmm_set_jigna <-  mrna_filter_jigna

pdata_tmm_jigna <- pdata_jigna
design <- cbind(model.matrix(~ 0 + pdata_tmm_jigna$Group))
colnames(design) <- gsub(".*\\$|)", "", colnames(design))

tmm_set_jigna <- DGEList(counts = tmm_set_jigna,
                   genes = mrna_gene_jigna[match(
                     row.names(mrna_filter_jigna), 
                     mrna_gene_jigna$gene_id),],
                   group = pdata_tmm_jigna$Group)

tmm_set_jigna <- calcNormFactors(tmm_set_jigna)

#--> Estimate dispersion and fit the model
tmm_set_jigna <- estimateGLMRobustDisp(tmm_set_jigna, design)

##### Count per million
tmm_cpm_jigna <- cpm(tmm_set_jigna, normalized.lib.sizes = TRUE, log = TRUE)

# --> Transpose the df to have the sample in lines
t_mrna_norm_jigna <- as.data.frame(t(tmm_cpm_jigna))
```
## RECOMBINE THE 2 DATAFRAME AFTER NORMALIZATION
```{r, message=FALSE}
X_combined_count_mrna <- rbind(t_mrna_norm_jigna,t_mrna_norm_stepniak)
Y_pdata <- read.csv("pdata.csv", sep =";", header = TRUE)

class <- as.factor(Y_pdata$Group)
study <- factor(Y_pdata$Study)
```
## Comparison with previous analysis
```{r, message=FALSE}
library(ggvenn)
genes_Jigna_downreg <- read.csv("../Differential_Analysis/Differential_analysis_Jigna/downregulated_genes_Jigna_samples.csv", header = TRUE)
genes_Jigna_upreg <- read.csv("../Differential_Analysis/Differential_analysis_Jigna/upregulated_genes_Jigna_samples.csv", header = TRUE)
#genes_Jigna_downreg_metabo <- read.csv("../Differential_Analysis/Differential_analysis_Jigna/downregulated_metabo_genes_Jigna_samples.csv", header = FALSE)
#genes_Jigna_upreg_metabbo <- read.csv("../Differential_Analysis/Differential_analysis_Jigna/upregulated_metabo_genes_Jigna_samples.csv", header = FALSE)

genes_Stepniak_downreg <- read.csv("../Differential_Analysis/Differential_Analysis_Stepniak/RNA/downregulated_genes_DA.csv", header = TRUE)
genes_Stepniak_upreg <- read.csv("../Differential_Analysis/Differential_Analysis_Stepniak/RNA/upregulated_genes_DA.csv", header = TRUE)
#genes_Stepniak_upreg_metabo <- read.csv("../Differential_Analysis/Differential_Analysis_Stepniak/RNA/upregulated_metabo_genes_DA.csv", header = TRUE)
#genes_Stepniak_downreg_metabo <- read.csv("../Differential_Analysis/Differential_Analysis_Stepniak/RNA/downregulated_metabo_genes_DA.csv", header = TRUE)

gene_names <- mrna_gene_stepniak[match(
  common_lines, 
  mrna_gene_stepniak$gene_id),]

metabo_genes_all <- read.table("../Differential_Analysis/Lists_genes/all_metabo.tsv", header = TRUE, sep = "\t")
metabo_genes <- metabo_genes_all$Gene_name

meta_analysis_genes <- gene_names$gene_name
down_genes_Jigna <- as.character(genes_Jigna_downreg$x)
up_genes_Jigna <- as.character(genes_Jigna_upreg$x)
genes_jigna <- c(down_genes_Jigna,up_genes_Jigna)

Jigna_genes = list(Downregulated = down_genes_Jigna, meta_analysis_genes = meta_analysis_genes, Metabo = metabo_genes, Upregulated = up_genes_Jigna)

# Create the Venn diagram
venn_plot_dysreg_jigna <- ggvenn(
  Jigna_genes,
)
venn_plot_dysreg_jigna

#######
meta_analysis_genes <- gene_names$gene_name
down_genes_Stepniak <- as.character(genes_Stepniak_downreg$x)
up_genes_Stepniak <- as.character(genes_Stepniak_upreg$x)
genes_stepniak <- c(down_genes_Stepniak,up_genes_Stepniak)

Stepniak_genes = list(Downregulated = down_genes_Stepniak, meta_analysis_genes = meta_analysis_genes, Metabo = metabo_genes, Upregulated = up_genes_Stepniak)

# Create the Venn diagram
venn_plot_dysreg_stepniak <- ggvenn(
  Stepniak_genes,
)
venn_plot_dysreg_stepniak

```
## MINT sPLS-DA Model
```{r, message=FALSE}
result <- mint.splsda(X = X_combined_count_mrna, Y = class, study = study, ncomp = 2)

class_colors <- c("#D60000", "#4294D8")

# Plot with custom colors
plotIndiv(result, comp = c(1, 2), col = class_colors, legend = TRUE)

## TUNING THE NUMBER OF COMPONENTS
splsda_perf <- perf(result)
plot(splsda_perf)
splsda_perf$global.error

optim_ncomp <- 2

## TUNING THE NUMBER OF FEATURES
splsda_tune <- tune(X = X_combined_count_mrna, 
                    Y = class, 
                    study = study, 
                    ncomp = optim_ncomp,
                    test.keepX = seq(1, 100, 1),
                    nrepeat = 5,
                    method = "mint.splsda",
                    measure = "BER",
                    dist = "centroids.dist",
                    )

plot(splsda_tune, sd = FALSE)

optim_keepX <- splsda_tune$choice.keepX # extract optimal values

optim_keepX <- c(500,500)

## FINAL MODEL
final_splsda_model <- mint.splsda(X = X_combined_count_mrna, 
                                  Y = class, 
                                  study = study, 
                                  ncomp = optim_ncomp, 
                                  keepX = optim_keepX)

#dev.set(dev.next())

#pdf("sim.pdf")

# Define a color palette for the class annotations
class_levels <- levels(as.factor(class))
class_colors <- c("#D60000", "#4294D8")
names(class_colors) <- class_levels

# Ensure row annotations are a factor
class <- as.factor(class)

row_colors <- class_colors[class]

legend <- list(legend = names(class_colors), fill = class_colors, 
       title = "Groups", cex = 0.8, inset = 0.02)

sim <- cim(final_splsda_model, comp = 1, margins=c(10,5), 
    row.sideColors = row_colors,
    legend = legend,
    row.names = FALSE, title = "MINT sPLS-DA, component 1",
    xlab = "Genes",
    ylab = "Samples"
    #,save = 'pdf', 
    #name.save = "Comp1_500genes2"
    )

#dev.off()

selected_gene_comp1 <- selectVar(final_splsda_model, comp = 1)$name

gene_names <- mrna_gene_stepniak[match(
  selected_gene_comp1, 
  mrna_gene_stepniak$gene_id),]


cluster_1 <- labels(sim$ddc[[1]])

gene_names_cluster_1 <- mrna_gene_stepniak[match(
  cluster_1, 
  mrna_gene_stepniak$gene_id),]

cluster_2 <- labels(sim$ddc[[2]])

gene_names_cluster_2 <- mrna_gene_stepniak[match(
  cluster_2, 
  mrna_gene_stepniak$gene_id),]

cluster1 <- gene_names_cluster_1$gene_name
cluster1
cluster2 <- gene_names_cluster_2$gene_name
cluster2

gene_names_cluster_1$cluster <- 1

gene_names_cluster_2$cluster <- 2

df <- rbind(gene_names_cluster_1, gene_names_cluster_2)

```
## COMPARISON WITH METABO
```{r, message = FALSE}
metabo_genes_all <- read.table("../Differential_Analysis/Lists_genes/all_metabo.tsv", header = TRUE, sep = "\t")
metabo_genes <- metabo_genes_all$Gene_name

gene_metabo <- intersect(metabo_genes,gene_names$gene_name)
gene_metabo

metabob_clust1 <- intersect(cluster1, metabo_genes)

metabo_clust2 <- intersect(cluster2, metabo_genes)
```
## Over representation analysis KEGG metabo
```{r, message = FALSE}
#---> Split genes according to the module they belong to
probes <- split(df$gene_name,
                f = df$cluster)

#--->  Over-representation analysis
library(ReactomePA)
library(clusterProfiler)
library(stringr)
library(stringi)
library(data.table)
library(scales)
library(plotly)

#--> Load annotations
library(gage)
library(org.Hs.eg.db)

#--> Load annotations
#gmt <- list.files("/shared/projects/gliomep53/Scripts_R/Differential_Analysis/gmt_file/", full.names = TRUE)

#--> Load KEGG annotation
kegg <- kegg.gsets(species = "hsa", id.type = "entrez")
 
#--> Select KEGG metabolic pathways
kegg.met <- kegg$kg.sets[kegg$met.idx]
 
#--> Map gene gene_name to Entrez ID
kegg.met_2 <- lapply(kegg.met, function(x) {mapIds(org.Hs.eg.db,x, 'SYMBOL', 'ENTREZID')})
names(kegg.met_2) <- names(kegg.met)

kegg.met_2 <- melt(kegg.met_2)
kegg.met_2 <- kegg.met_2[,c(2,1)]
colnames(kegg.met_2) <- c("term", "gene")

#--> Create a formatting function for gene names
format <- function(x) {
  
  b <- gregexpr("/", x)[[1]]
  
  if (length(b) > 10) {
    stri_sub_all(x, from = b[seq(from = 10, to = length(b), by = 10)], length = 1) <- " \n "
    x <- gsub("/", " ", x)
    
  }else{
    x <- gsub("/", " ", x)
  }
  
  return(x)
}

#--> Create a function for over-representation analysis
or_analysis <- function(gmt) {
  
  #gmt <- read.gmt(gmt)
  gmt <- gmt
  
  fun_enrich_res <- data.frame(compareCluster(probes,
                                              fun='enricher',
                                              TERM2GENE=gmt,
                                              pvalueCutoff =1,
                                              qvalueCutoff =1,
                                              minGSSize = 10))
  
  sel <- tapply(fun_enrich_res$Description,
                INDEX = fun_enrich_res$Cluster,
                function(x) {y <- x[c(1:10)]
                return(y)})
  sel <- unlist(sel)
  
  fun_enrich_res_2 <- fun_enrich_res[fun_enrich_res$Description %in%sel,]
  
  for (i in sel) {
    if (nrow(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]) > 1) {
      min_p <- min(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]$pvalue)
      fun_enrich_res_2[which(fun_enrich_res_2$pvalue > min_p &
                               fun_enrich_res_2$Description == i),]$qvalue <- NA
    }}
  
  #fun_enrich_res_2 <- fun_enrich_res[which(fun_enrich_res$p.adjust < .1),]
  
  #--> Format names
  fun_enrich_res_2 <- na.omit(fun_enrich_res_2)
  
  #--> Order
  fun_enrich_res_2$Description <- factor(fun_enrich_res_2$Description,
                                         levels = unique(fun_enrich_res_2$Description))
  
  #--> Convert to gene ratios to percent
  fun_enrich_res_2$Percent <- unlist(lapply(strsplit(fun_enrich_res_2$GeneRatio, "/"), function(x){as.numeric(x[1])/as.numeric(x[2])}*100))
  
  #--> Format gene names
  fun_enrich_res_2$geneID <- lapply(fun_enrich_res_2$geneID, format)
  
  return(list(Res = fun_enrich_res_2,
              Levels = levels(fun_enrich_res_2$Description)))
}

#--> Apply over-representation tests
test <- lapply(list(kegg.met_2), or_analysis)

#--> Format the output
test_2 <- list(test[[1]]$Res)

names(test_2) <- c("KEGG")

test_2 <- rbindlist(test_2, use.names = TRUE, idcol = "Database")

test_2$Description <- as.factor(test_2$Description)

levels(test_2$Description) <- c(test[[1]]$Levels)

test_2$Description <- gsub("KEGG_|", "", test_2$Description)
test_2$Description <- gsub("_", " ", test_2$Description)
test_2$Description <- str_to_sentence(test_2$Description)
test_2$Description <- factor(test_2$Description, levels = unique(test_2$Description))

#--> Plot
p <- ggplot(test_2, aes(y = Description,
                        x = Cluster,
                        color = -log10(qvalue),
                        size = Percent)) +
  geom_point(shape = 16, stroke = .5, aes(text = gsub("/", " ", geneID))) +
  scale_color_gradient(low = "firebrick1", high = "chartreuse3") +
  facet_wrap(vars(Database), scales = "free_y", ncol = 1) +
  theme() + 
  scale_y_discrete(labels = label_wrap(70)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.margin = margin(10, 10, 10, 200))

ggplotly(p)
```
###### Over representation analysis KEGG
```{r, message=FALSE}
#---> Split genes according to the module they belong to
probes <- split(df$gene_name,
                f = df$cluster)

#--->  Over-representation analysis
library(ReactomePA)
library(clusterProfiler)
library(stringr)
library(stringi)
library(data.table)
library(scales)

#--> Load annotations
gmt <- list.files("/shared/projects/gliomep53/Scripts_R/Differential_Analysis/gmt_file/kegg/", full.names = TRUE)

#--> Create a formatting function for gene names
format <- function(x) {
  
  b <- gregexpr("/", x)[[1]]
  
  if (length(b) > 10) {
    stri_sub_all(x, from = b[seq(from = 10, to = length(b), by = 10)], length = 1) <- " \n "
    x <- gsub("/", " ", x)
    
  }else{
    x <- gsub("/", " ", x)
  }
  
  return(x)
}

#--> Create a function for over-representation analysis
or_analysis <- function(gmt) {
  
  gmt <- read.gmt(gmt)
  
  fun_enrich_res <- data.frame(compareCluster(probes,
                                              fun='enricher',
                                              TERM2GENE=gmt,
                                              pvalueCutoff =1,
                                              qvalueCutoff =.2,
                                              minGSSize = 10))
  
  sel <- tapply(fun_enrich_res$Description,
                INDEX = fun_enrich_res$Cluster,
                function(x) {y <- x[c(1:5)]
                return(y)})
  sel <- unlist(sel)
  
  fun_enrich_res_2 <- fun_enrich_res[fun_enrich_res$Description %in%sel,]
  
  for (i in sel) {
    if (nrow(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]) > 1) {
      min_p <- min(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]$pvalue)
      fun_enrich_res_2[which(fun_enrich_res_2$pvalue > min_p &
                               fun_enrich_res_2$Description == i),]$qvalue <- NA
    }}
  
  #fun_enrich_res_2 <- fun_enrich_res[which(fun_enrich_res$p.adjust < .1),]
  
  #--> Format names
  fun_enrich_res_2 <- na.omit(fun_enrich_res_2)
  
  #--> Order
  fun_enrich_res_2$Description <- factor(fun_enrich_res_2$Description,
                                         levels = unique(fun_enrich_res_2$Description))
  
  #--> Convert to gene ratios to percent
  fun_enrich_res_2$Percent <- unlist(lapply(strsplit(fun_enrich_res_2$GeneRatio, "/"), function(x){as.numeric(x[1])/as.numeric(x[2])}*100))
  
  #--> Format gene names
  fun_enrich_res_2$geneID <- lapply(fun_enrich_res_2$geneID, format)
  
  return(list(Res = fun_enrich_res_2,
              Levels = levels(fun_enrich_res_2$Description)))
}

#--> Apply over-representation tests
test <- lapply(gmt, or_analysis)

#--> Format the output
test_2 <- list(test[[1]]$Res)

names(test_2) <- c("KEGG")

test_2 <- rbindlist(test_2, use.names = TRUE, idcol = "Database")

test_2$Description <- as.factor(test_2$Description)

levels(test_2$Description) <- c(test[[1]]$Levels)

test_2$Description <- gsub("KEGG_|", "", test_2$Description)
test_2$Description <- gsub("_", " ", test_2$Description)
test_2$Description <- str_to_sentence(test_2$Description)
test_2$Description <- factor(test_2$Description, levels = unique(test_2$Description))

#--> Plot
p <- ggplot(test_2, aes(y = Description,
                        x = Cluster,
                        color = -log10(qvalue),
                        size = Percent)) +
  geom_point(shape = 16, stroke = .5, aes(text = gsub("/", " ", geneID))) +
  scale_color_gradient(low = "firebrick1", high = "chartreuse3") +
  facet_wrap(vars(Database), scales = "free_y", ncol = 1) +
  theme() + 
  scale_y_discrete(labels = label_wrap(70)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.margin = margin(10, 10, 10, 200))

ggplotly(p)
```
###### Over representation analysis Reactome
```{r, message=FALSE}
#---> Split genes according to the module they belong to
probes <- split(df$gene_name,
                f = df$cluster)

#--->  Over-representation analysis
library(ReactomePA)
library(clusterProfiler)
library(stringr)
library(stringi)
library(data.table)
library(scales)

#--> Load annotations
gmt <- list.files("/shared/projects/gliomep53/Scripts_R/Differential_Analysis/gmt_file/reactome/", full.names = TRUE)

#gmt <- read.gmt("/shared/projects/gliomep53/Scripts_R/Differential_Analysis/gmt_file/reactomev2023.txt")

#--> Create a formatting function for gene names
format <- function(x) {
  
  b <- gregexpr("/", x)[[1]]
  
  if (length(b) > 10) {
    stri_sub_all(x, from = b[seq(from = 10, to = length(b), by = 10)], length = 1) <- " \n "
    x <- gsub("/", " ", x)
    
  }else{
    x <- gsub("/", " ", x)
  }
  
  return(x)
}

#--> Create a function for over-representation analysis
or_analysis <- function(gmt) {
  
  gmt <- read.gmt(gmt)
  
  fun_enrich_res <- data.frame(compareCluster(probes,
                                              fun='enricher',
                                              TERM2GENE=gmt,
                                              pvalueCutoff =1,
                                              qvalueCutoff =.2,
                                              minGSSize = 10))
  
  sel <- tapply(fun_enrich_res$Description,
                INDEX = fun_enrich_res$Cluster,
                function(x) {y <- x[c(1:5)]
                return(y)})
  sel <- unlist(sel)
  
  fun_enrich_res_2 <- fun_enrich_res[fun_enrich_res$Description %in%sel,]
  
  for (i in sel) {
    if (nrow(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]) > 1) {
      min_p <- min(fun_enrich_res_2[which(fun_enrich_res_2$Description == i),]$pvalue)
      fun_enrich_res_2[which(fun_enrich_res_2$pvalue > min_p &
                               fun_enrich_res_2$Description == i),]$qvalue <- NA
    }}
  
  #fun_enrich_res_2 <- fun_enrich_res[which(fun_enrich_res$p.adjust < .1),]
  
  #--> Format names
  fun_enrich_res_2 <- na.omit(fun_enrich_res_2)
  
  #--> Order
  fun_enrich_res_2$Description <- factor(fun_enrich_res_2$Description,
                                         levels = unique(fun_enrich_res_2$Description))
  
  #--> Convert to gene ratios to percent
  fun_enrich_res_2$Percent <- unlist(lapply(strsplit(fun_enrich_res_2$GeneRatio, "/"), function(x){as.numeric(x[1])/as.numeric(x[2])}*100))
  
  #--> Format gene names
  fun_enrich_res_2$geneID <- lapply(fun_enrich_res_2$geneID, format)
  
  return(list(Res = fun_enrich_res_2,
              Levels = levels(fun_enrich_res_2$Description)))
}

#--> Apply over-representation tests
test <- lapply(gmt, or_analysis)

#--> Format the output
test_2 <- list(test[[1]]$Res)

names(test_2) <- c("Reactome")

test_2 <- rbindlist(test_2, use.names = TRUE, idcol = "Database")

test_2$Description <- as.factor(test_2$Description)

levels(test_2$Description) <- c(test[[1]]$Levels)

test_2$Description <- gsub("REACTOME_|", "", test_2$Description)
test_2$Description <- gsub("_", " ", test_2$Description)
test_2$Description <- str_to_sentence(test_2$Description)
test_2$Description <- factor(test_2$Description, levels = unique(test_2$Description))

#--> Plot
p <- ggplot(test_2, aes(y = Description,
                        x = Cluster,
                        color = -log10(qvalue),
                        size = Percent)) +
  geom_point(shape = 16, stroke = .5, aes(text = gsub("/", " ", geneID))) +
  scale_color_gradient(low = "firebrick1", high = "chartreuse3") +
  facet_wrap(vars(Database), scales = "free_y", ncol = 1) +
  theme() + 
  scale_y_discrete(labels = label_wrap(70)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.margin = margin(10, 10, 10, 200))

ggplotly(p)
```