---
title: "Differential_Analysis_Stepniak_ChiP_H3K27ac_0,2"
author: "Alizée LANON"
date: "2024-05-06"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,     # Suppress warnings from R code
  message = FALSE     
)
```
###Process of the data
The raw data were processed using nf-core v2.14.1 chipseq v2.0.0 pipeline on the IFB cluster with the following command line:
"
sbatch --wrap="\
nextflow run nf-core/chipseq -profile ifb_core -resume -r 2.0.0\
                --input ./Glioma/Stepniak/ChiPseq/samplesheet_ChIPseq_2.csv \
                --outdir ./Glioma/Stepniak/ChiPseq/nextflow2/ \
                --fasta ./Glioma/save/RESSOURCES/REFERENCES/GRCh38/Homo_sapiens.GRCh38.dna_sm.primary_assembly.fa \
                --gtf ./Glioma/save/RESSOURCES/REFERENCES/GRCh38/Homo_sapiens.GRCh38.108.gtf \
                --bwa_index ./Glioma/save/RESSOURCES/REFERENCES/GRCh38/bwa \
                --blacklist ./Glioma/save/RESSOURCES/REFERENCES/GRCh38/hg38-blacklist.v2.bed \
                --save_unaligned \
                --narrow_peak \
                --macs_gsize 2652783500"
"
### Import 
```{r, message=FALSE}
library(ggplot2)
library(BiocStyle)

#--> Phenotypic matrix 
pdata <- read.csv("pdata/pdata_chip_H3K27ac.csv", sep =";", header = TRUE)
#pdata <- pdata[-6,]
row.names(pdata) <- pdata$Sample

#--> Count matrix
# /shared/projects/gliomep53/Scripts_R/Differential_Analysis_Stepniak/CHiP/
mrna_raw_counts <- read.delim(file = "../../../../../Glioma/Stepniak/ChiPseq/Count_matrix/H3K27ac.consensus_peaks.featureCounts.txt", header = TRUE, sep = "\t")
row.names(mrna_raw_counts) <- mrna_raw_counts$Geneid
mrna_raw_counts_table <- mrna_raw_counts[,c(7,8,10,11,12,13)]

#--> Order and match
mrna_raw_counts_table <- mrna_raw_counts[,match(pdata$Sample, colnames(mrna_raw_counts), nomatch = 0)]
pdata <- pdata[pdata$Sample %in% colnames(mrna_raw_counts),]

mrna_raw_counts_info <- mrna_raw_counts[,c(1,2,3,4,5)]

#--> Feature annotation
mrna_gene <- read.table("../../../../../Glioma/Stepniak/ChiPseq/Count_matrix/H3K27ac.consensus_peaks.featureCounts.txt", header = TRUE)[,c(1)]
```
### Filter 
The filtering of the genes is the step that allows to filter out the genes that are not expressed in any of the conditions. To do so, the package HTSFilter (v1.24.0) was used. It takes as an input the raw count matrix as well as an object in which each samples is associated at its experimental group (IDHwt or IDHmut). As we are in ChiPseq data, this step is not necessary because this method only detects accessible chromatin area.
```{r, message=FALSE}
#--> Filter lowly expressed genes
library(HTSFilter)

mrna_filter <- HTSFilter(round(mrna_raw_counts_table),
                         pdata$Group,
                         s.min=1, s.max=200)
```
### Normalization 
The forst normalization is done with the EDAseq package (v2.28.0) to conduct a "between lane normalization" (to look at sequencing depth and other distributional differences between lanes). Then edgeR package (v1.36.0) is used to perform the differential analysis. First, a DGEList object is created, we apply to this object a normalization by the library size. As edgeR based its analysis on the hypothesis of a negative binomial dispersion of the genes expression, the function "estimateGLMRobustDisp" is applied to perform a robust estimation of the dispersion parameter for each gene.
```{r, message=FALSE}
library(RUVSeq)
library(reshape)
library(EDASeq)

#--> Quantile normalization
raw_set <- newSeqExpressionSet(counts = as.matrix(mrna_filter$filteredData))
colnames(raw_set) <- pdata$Sample
qn_set <- betweenLaneNormalization(raw_set, which = "full")

#--> edgeR TMM normalization
library(edgeR)

tmm_set <-  mrna_filter$filteredData

pdata_tmm <- pdata
design <- cbind(model.matrix(~ 0 + pdata_tmm$Group))
colnames(design) <- gsub(".*\\$|)", "", colnames(design))

tmm_set <- DGEList(counts = tmm_set,
                   genes = mrna_gene,
                   group = pdata_tmm$Group)

tmm_set <- calcNormFactors(tmm_set)

#--> Estimate dispersion and fit the model
tmm_set <- estimateGLMRobustDisp(tmm_set, design)
```
### Quality Check
The quality check is perform to look at the distributio of gene expression after normalization. It allows to see if a sample is an outlier and if so, which normalization corrects "better" this outlier to align with other samples.
```{r, message=FALSE}
##### Boxplot
library(plotly)
#--> Get data
bp_full <- rbind(melt(log2(mrna_filter$filteredData + 1)),
                 melt(log2(normCounts(qn_set) + 1)),
                 melt(cpm(tmm_set, normalized.lib.sizes = TRUE, log = TRUE)))

#--> Build the data.frame
bp_full$Method <- rep(c("Before QN", "After QN", "After TMM"),
                      each = nrow(mrna_filter$filteredData) *    ncol(mrna_filter$filteredData))

bp_full$Method <- factor(bp_full$Method, levels = c("Before QN",
                                                    "After QN",
                                                    "After TMM"))

colnames(bp_full) <- c("Gene", "Sample", "Expression", "Method")

bp_full <- merge(bp_full, pdata,
                 by.x = "Sample", by.y = "Sample",
                 all = TRUE)

#--> Compute statistics
calc_stat <- function(x) {
  coef <- 1.5
  n <- sum(!is.na(x))
  stats <- quantile(x, probs = c(0.1, 0.25, 0.5, 0.75, 0.9))
  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
  return(stats)
}

#--> Build the plot                 
bp <- ggplot(bp_full, aes(x = Sample, y = Expression, fill = Group, text = Sample)) + 
  stat_summary(fun.data = calc_stat, geom = "boxplot") +
  facet_wrap(~Method, scales = "free_y") +
  ggtitle("Distribution of gene expression")  +
  ylab("Expression (logCPM)") + xlab("") +
  theme() +  # Call the theme function with parentheses
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")


ggplotly(bp)

```
### PCA prcomp
Then, the factoextra package (v1.0.7), dependencie of FactoMineR package (v2.4) was used to perform Principal Component Analysis, an unsupervised clustering method. It was performed with two differents functions, you can chose and only keep one for further analysis.
```{r, message=FALSE}
library(ggrepel)
library(FactoMineR)
library(factoextra)
#-->  Creating the expression matrices
data_pca_before_qn <- log2(mrna_filter$filteredData + 1)
data_pca_after_qn  <- log2(normCounts(qn_set) + 1)
data_pca_after_tmm <- cpm(tmm_set, normalized.lib.sizes = TRUE, log = TRUE)

#-->  Computing PCA
pca_before_qn   <- as.data.frame(prcomp(t(na.omit(data_pca_before_qn)), scale. = TRUE, center = TRUE)$x)

pca_after_qn    <- as.data.frame(prcomp(t(na.omit(data_pca_after_qn)), scale. = FALSE, center = TRUE)$x) 

pca_after_tmm   <- as.data.frame(prcomp(t(na.omit(data_pca_after_tmm)), scale. = TRUE, center = TRUE)$x)

#-->  Combine
comb <- data.frame(rbind(
  pca_before_qn[,c(1:2)],
  pca_after_qn[,c(1:2)],
  pca_after_tmm[,c(1:2)]),
  PCA = rep(c("Before QN", "After QN",
              "After TMM"), each = nrow(pdata)),
  Sample = pdata$Sample,
  Group = pdata$Group)
colnames(comb) <- c("PC1", "PC2", "PCA", "Sample","Group")

comb$PCA <- factor(comb$PCA, levels = c("Before QN", "After QN",
                                        "After TMM"))

#-->  Plot
pca_plot <- ggplot(comb, aes(x= PC1, y = PC2)) + 
  geom_point(aes(color = Group,
                 text = Sample)) +
  facet_wrap(~PCA, nrow = 2, scales = "free") +
  ggtitle("PCA - Normalization")  +
  theme()

ggplotly(pca_plot)
```
### PCA method 2
```{r, message=FALSE}
library(ggrepel)
library(FactoMineR)
library(factoextra)
#-->  Creating the expression matrices
data_pca_before_qn <- log2(mrna_filter$filteredData + 1)
data_pca_after_qn  <- log2(normCounts(qn_set) + 1)
data_pca_after_tmm <- cpm(tmm_set, normalized.lib.sizes = TRUE, log = TRUE)

#-->  Computing PCA
res_data_pca_before_qn <- PCA(t(na.omit(data_pca_before_qn)), scale.unit = TRUE,graph = FALSE)

plot_data_pca_before_qn <- fviz_pca_ind(res_data_pca_before_qn, title = "Pca of DA samples before qn",
             col.ind = pdata$Group, # color by groups
             palette = c("#D60000", "#4294D8"),
             legend.title = "Groups",
             ylim = c(-200, 300), xlim = c(-350,300))

res_data_pca_after_qn <- PCA(t(na.omit(data_pca_after_qn)), scale.unit = TRUE,graph = FALSE)

plot_data_pca_after_qn <- fviz_pca_ind(res_data_pca_after_qn, title = "Pca of DA samples after qn",
             col.ind = pdata$Group, # color by groups
             palette = c("#D60000", "#4294D8"),
             legend.title = "Groups",
             ylim = c(-300, 300), xlim = c(-400,300))

res_data_pca_after_tmm <- PCA(t(na.omit(data_pca_after_tmm)), scale.unit = TRUE,graph = FALSE)

plot_data_pca_after_tmm <- fviz_pca_ind(res_data_pca_after_tmm, title = "Pca of DA samples after tmm",
             col.ind = pdata$Group, # color by groups
             palette = c("#D60000", "#4294D8"),
             legend.title = "Groups",
             ylim = c(-350, 300), xlim = c(-500,300))

## Plot 
par(mfrow = c(1, 3))
print(plot_data_pca_before_qn)
print(plot_data_pca_after_qn)
print(plot_data_pca_after_tmm)

```
### Differential level of modification analysis
Then, the "glmFit" function is used to fit a generalized linear model to the count data. Afterwards, based on the fitted GLM model, the function "glmLRT" is used to conduct a likelihood ratio test to identify differentially expressed genes. Next, we extract the top differentially expressed genes, we then adjust their p-values for multiple testing with the method "False Discovery Rate" and we filter out the genes with an adjusted p-value above 0.1
```{r, message=FALSE}
### ANOVA like test
#--> Estimate dispersion and fit the model
tmm_fit <- glmFit(tmm_set, design = design)

################################################
#--> ANOVA-like test
lrt_full <- glmLRT(tmm_fit,
                   contrast =  makeContrasts(IDHmut_vs_IDHwt = GroupIDHmut - GroupIDHwt,
                                             levels = design))

#--> List DE genes and correct for multiple testing
top_full <- topTags(lrt_full, n = Inf)$table
top <- topTags(lrt_full, adjust.method = "fdr",
               p.value = 0.2,  n = Inf)$table

# Extract fold change values and gene names
fold_changes <- top$logFC
gene_names <- top$genes

# Identify up-regulated and down-regulated genes
upregulated_loci <- gene_names[fold_changes > 0]
up_loci_Grange <- mrna_raw_counts_info[upregulated_loci,]

downregulated_loci <- gene_names[fold_changes < 0]
down_loci_Grange <- mrna_raw_counts_info[downregulated_loci,]

```
### EXTRACT THE 5000 LOCI WITH HIGHEST LOGfc FOR INTEGRATIVE ANALYSIS 
Not a step of differential analysis but only to extract the 5000 loci (2500 positive logFC and 2500 negative logFC) for an integrative analysis.
```{r, message = FALSE}
library(dplyr)
top_complete <- top_full
#sorted_genes_top <- top_complete[order(top_complete$logFC), ]

sorted_genes_df_desc <- top_complete %>%
  arrange(desc(logFC))

sorted_genes_df_asc <- top_complete %>%
  arrange(logFC)

sorted_genes_df_desc_top <- sorted_genes_df_desc[1:2500,]
sorted_genes_df_asc_top <- sorted_genes_df_asc[1:2500,]

sorted_genes_df_top <- rbind(sorted_genes_df_desc_top, sorted_genes_df_asc_top)
#write.table(sorted_genes_df_top, "../../../Integrative_Analysis/Data/5000_peaks_logFC_H3K27ac_Stepniak.csv", row.names = FALSE)
```
### Hierarchical clustering
In order to vizualize those results, a heatmap was chosen as a representation. This representation,in addition to allowing the vizualization of the clusters of dysregulated loci among the samples, it also allows the display of clusterized samples. To do so, we perform upstream a k-means clustering (or hierarchical clustering) with the factoextra and cluster (v2.1.2) packages and the "ward.D2" method. Finally, the heatmap is created with the package pheatmap (v1.0.12) and the function of the same name.
```{r, message=FALSE}
library(pheatmap)
library(viridis)

#--> Select ID of DE genes
sel_hm <- top$genes

#--> Get expression matrix
counts_mrna <- edgeR::cpm(tmm_set, normalized.lib.sizes = TRUE, log = TRUE)
sel_hm_mrna <- counts_mrna[match(sel_hm, row.names(counts_mrna), nomatch = 0),]

#write.table(counts_mrna,"../../../Integrative_Analysis/Data/counts_ChIP_H3K27ac_Stepniak_norm.csv", row.names = TRUE)

library(cluster)
library(factoextra)

#create plot of number of clusters vs total within sum of squares
#fviz_nbclust(counts_mrna, kmeans, method = "wss")

#--> Perform K-means clustering
clustering_hc <-  hclust(as.dist(1-cor(t(sel_hm_mrna))), method = "ward.D2")
clusters <- cutree(tree = clustering_hc , k = 2)

#--> Order genes
sel_hm_mrna <- sel_hm_mrna[clustering_hc$order,]

#--> Create row annotation matrix
annot_row = data.frame(row.names = row.names(sel_hm_mrna),
                       cluster = as.factor(clusters[clustering_hc$order]))

#--> Create column annotation matrix
pdata_chip <- pdata
colnames(sel_hm_mrna) <- row.names(pdata_chip)

# Annotation Colors
annotation_colors <- list(
  Group = c("IDHmut" = "#D60000", "IDHwt" = "#4294D8")    # Colors for column annotations
)

#--> Build the heatmap
pheatmap(sel_hm_mrna,
         scale = "row",
         cluster_rows = FALSE,
         cluster_cols = TRUE, show_rownames = FALSE,
         clustering_method  = "ward.D2",
         clustering_distance_cols = "correlation",
         annotation_row = annot_row["cluster"],
         annotation_col = pdata[2],
         main = "Clustering ",
         color = viridis(100),
         labels_row = NA,
         fontsize = 6,
         annotation_colors = annotation_colors)
```
### Table
```{r, message=FALSE}
#library(DT)

#table <- merge(top, annot_row, by = "row.names")

#datatable(table[,-c(1,12,13)], extensions = "Buttons",
#          options = list(buttons = c("csv", "excel"))) %>%
#  formatRound(columns = c(3:9), digits=2)
```
### Peak Annotation
Peak annotation is the step where for each loci we attribute the name of a gene. Several methods exists bbut here we used the nearest gene method.
I used ChIPpeakAnno package (v3.28.1), to attribute to each loci the name of the closest gene.
```{r, message=FALSE}
library(ChIPpeakAnno)
library(EnsDb.Hsapiens.v86)
library(org.Hs.eg.db)
library(biomaRt)

gr_upreg <- GRanges(
    seqnames = Rle(up_loci_Grange$Chr),
    ranges = IRanges(start = up_loci_Grange$Start, end = up_loci_Grange$End),
    strand = Rle(strand(up_loci_Grange$Strand)))

gr_downreg <- GRanges(
    seqnames = Rle(down_loci_Grange$Chr),
    ranges = IRanges(start = down_loci_Grange$Start, end = down_loci_Grange$End),
    strand = Rle(strand(down_loci_Grange$Strand)))

ensembl_hs86_transcript <- transcripts(EnsDb.Hsapiens.v86)

macs_peak_ensembl_upreg <- annotatePeakInBatch(gr_upreg, 
                                         AnnotationData = ensembl_hs86_transcript)

macs_peak_ensembl_downreg <- annotatePeakInBatch(gr_downreg, 
                                         AnnotationData = ensembl_hs86_transcript)


mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL",
                dataset = "hsapiens_gene_ensembl")

macs_peak_ensembl_upregID <- addGeneIDs(annotatedPeak = macs_peak_ensembl_upreg, 
                                mart = mart,
                                feature_id_type = "ensembl_transcript_id",
                                IDs2Add = "hgnc_symbol")

macs_peak_ensembl_downregID <- addGeneIDs(annotatedPeak = macs_peak_ensembl_downreg, 
                                mart = mart,
                                feature_id_type = "ensembl_transcript_id",
                                IDs2Add = "hgnc_symbol")

genes_upreg <- macs_peak_ensembl_upregID$hgnc_symbol

genes_downreg <- macs_peak_ensembl_downregID$hgnc_symbol

metabo_genes_all <- read.table("../../Lists_genes/all_metabo.tsv", header = TRUE, sep = "\t")
metabo_genes <- metabo_genes_all$Gene_name

metabo_up <- intersect(metabo_genes,genes_upreg)
metabo_down <- intersect(metabo_genes,genes_downreg)

genes_upreg
genes_downreg
metabo_up
metabo_down
```
### Comparison with differential analysis
We compared the results obtaines with differential analysis results
```{r, message=FALSE}
genes_Stepniak_upreg_metabo <- read.csv("../RNA/lists_dysreg_genes_stepniak/upregulated_metabo_genes_DA.csv", header = TRUE)
genes_Stepniak_downreg_metabo <- read.csv("../RNA/lists_dysreg_genes_stepniak/downregulated_metabo_genes_DA.csv", header = TRUE)

genes_Jigna_downreg_metabo <- read.csv("../../Differential_Analysis_Jigna/lists_genes_up_down_jigna_data/downregulated_metabo_genes_Jigna_samples.csv", header = TRUE)
genes_Jigna_upreg_metabo <- read.csv("../../Differential_Analysis_Jigna/lists_genes_up_down_jigna_data/upregulated_metabo_genes_Jigna_samples.csv", header = TRUE)

list_genes_diff_analysis_1 <- append(genes_Jigna_downreg_metabo$x, genes_Jigna_upreg_metabo$x)
list_genes_diff_analysis_2 <- append(genes_Stepniak_downreg_metabo$x, genes_Stepniak_upreg_metabo$x)
list_genes_diff_analysis_1 <- unlist(list_genes_diff_analysis_1)
list_genes_diff_analysis_2 <- unlist(list_genes_diff_analysis_2)

list_genes_diff_analysis <- append(list_genes_diff_analysis_1,list_genes_diff_analysis_2)


loci_upreg <- intersect(list_genes_diff_analysis, genes_upreg)
loci_downreg <- intersect(list_genes_diff_analysis,genes_downreg)

```
### Comparison with H3K4me3
We compare with H3K4me3 because a combination between those two marks can signify an activation or repression of a promotor depending on the combination of the marks.
```{r, message = FALSE}
list_genes_upreg_H3K4me3 <- read.csv("Lists_genes_locis_dysreg/genes_upreg_H3K4me3.csv", header = TRUE)
list_genes_downreg_H3K4me3 <- read.csv("Lists_genes_locis_dysreg/genes_downreg_H3K4me3.csv", header = TRUE)
list_metabo_upreg_H3K4me3 <- read.csv("Lists_genes_locis_dysreg/genes_metabo_upreg_H3K4me3.csv", header = TRUE)
list_metabo_downreg_H3K4me3 <- read.csv("Lists_genes_locis_dysreg/genes_metabo_downreg_H3K4me3.csv", header = TRUE)

genes_H3K4me3up_H3K27acup <- intersect(list_genes_upreg_H3K4me3$x,genes_upreg)
genes_H3K4me3down_H3K27acdown <- intersect(list_genes_downreg_H3K4me3$x,genes_downreg)
#genes_H3K4me3up_H3K27acdown <- intersect(list_genes_upreg_H3K4me3$x,genes_downreg)
#genes_H3K4me3down_H3K27acup <- intersect(list_genes_downreg_H3K4me3$x,genes_upreg)

metabo_up_between_2_marks <- intersect(metabo_genes, genes_H3K4me3up_H3K27acup)
metabo_down_between_2_marks <- intersect(metabo_genes, genes_H3K4me3down_H3K27acdown)

genes_H3K4me3up_H3K27acup
genes_H3K4me3down_H3K27acdown
metabo_up_between_2_marks
metabo_down_between_2_marks
```
### Enrichment Analysis GREAT
The last analysis performed here is an enrichment analysis using GREAT.
It allows to know in which pathway are enriched the information given by the modified histons marks.
```{r, message = FALSE}
library(rGREAT)

######## UP
new_seqlevels_up <- paste0("chr", seqlevels(gr_upreg))

seqlevels(gr_upreg) <- new_seqlevels_up

# Now update the seqnames with the "chr" prefix
new_seqnames_up <- paste0("chr", seqnames(gr_upreg))

GREAT_up <- submitGreatJob(gr_upreg, 
    bg = NULL,
    species               = "hg38",
    includeCuratedRegDoms = TRUE,
    rule                  = c("basalPlusExt"),
    adv_upstream          = 5.0,
    adv_downstream        = 1.0,
    adv_span              = 1000.0,
    request_interval = 60,
    max_tries = 10,
    version =  "4.0.4",
    base_url = "http://great.stanford.edu/public/cgi-bin",
    help = TRUE)

enrichment_table_up <- getEnrichmentTables(GREAT_up)

#enrichment_table_up$`GO Molecular Function`
#enrichment_table_up$`GO Biological Process`
#enrichment_table_up$`GO Cellular Component`

#write.table(enrichment_table_up$`GO Molecular Function`, "Enrich_table_GO_H3K27ac/enrichment_table_up_H3K27ac_GO_molecular_function.csv", row.names = FALSE, sep =";" )
#write.table(enrichment_table_up$`GO Biological Process`, "Enrich_table_GO_H3K27ac/enrichment_table_up_H3K27ac_GO_biological_process.csv", row.names = FALSE, sep =";" )
#write.table(enrichment_table_up$`GO Cellular Component`, "Enrich_table_GO_H3K27ac/enrichment_table_up_H3K27ac_GO_cellular_component.csv", row.names = FALSE, sep =";" )

######## DOWN
new_seqlevels_down <- paste0("chr", seqlevels(gr_downreg))

seqlevels(gr_downreg) <- new_seqlevels_down

# Now update the seqnames with the "chr" prefix
new_seqnames_down <- paste0("chr", seqnames(gr_downreg))


GREAT_down <- submitGreatJob(gr_downreg, 
    bg = NULL,
    species               = "hg38",
    includeCuratedRegDoms = TRUE,
    rule                  = c("basalPlusExt"),
    adv_upstream          = 5.0,
    adv_downstream        = 1.0,
    adv_span              = 1000.0,
    request_interval = 60,
    max_tries = 10,
    version =  "4.0.4",
    base_url = "http://great.stanford.edu/public/cgi-bin",
    help = TRUE)

enrichment_table_down <- getEnrichmentTables(GREAT_down)
#enrichment_table_down$`GO Molecular Function`
#enrichment_table_down$`GO Biological Process`
#enrichment_table_down$`GO Cellular Component`

#write.table(enrichment_table_down$`GO Molecular Function`, "Enrich_table_GO_H3K27ac/enrichment_table_down_H3K27ac_GO_molecular_function.csv", row.names = FALSE, sep =";" )
#write.table(enrichment_table_down$`GO Biological Process`, "Enrich_table_GO_H3K27ac/enrichment_table_down_H3K27ac_GO_biological_process.csv", row.names = FALSE, sep =";" )
#write.table(enrichment_table_down$`GO Cellular Component`, "Enrich_table_GO_H3K27ac/enrichment_table_down_H3K27ac_GO_cellular_component.csv", row.names = FALSE, sep =";" )

```